---
title: 'Integration'
author: 'Christoph Hafemeister'
date: '`r format(Sys.time(), "%B %d, %Y %H:%M:%S %Z")`'
output:
  html_document:
    toc: TRUE
    code_folding: "hide"
params:
  config_file: "config.yaml"
---

```{r}
# load project-specific parameters
config <- yaml::read_yaml(params$config_file)

suppressPackageStartupMessages({
  library('Matrix')
  library('dplyr')
  library('Seurat')
  library('sctransform')
  library('reshape2')
  library('ggplot2')
  library('patchwork')
  source(file.path(config$project_root, 'R', 'utils.R'))
})

future::plan(strategy = "sequential")
options(future.globals.maxSize = 8 * 1024 ^ 3)
#options(future.fork.enable = TRUE)

knitr::opts_chunk$set(comment = NA, fig.width = 5, fig.height = 3.5, out.width = '60%',
                      warning = TRUE, error = TRUE, echo = TRUE, message = TRUE,
                      dpi = 100)
options(dplyr.summarise.inform = FALSE)
options(tibble.width = Inf)
options(DT.options = list(paging = FALSE, scrollX = TRUE, scrollY = '350px',
                          scrollCollapse = TRUE, searching = FALSE, autoWidth = TRUE))

old <- theme_set(theme_bw(base_size = 11))
set.seed(8569205)
start_time <- proc.time()['elapsed']

rds_dir <- file.path(config$out_root, 'Rds')
```

Read merged object; add meta data; subset to samples of interest
```{r}
s <- readRDS(file = file.path(rds_dir, '01_s_merged.rds'))
s$genotype <- factor(s$orig.ident)
levels(s$genotype) <- c('WT', 'WT', 'WT', 'WT', 
                        'LTBR mutant', 'LTBR mutant')

print(table(s$orig.ident, s$genotype))
```

Perform ["standard" Seurat integration](https://satijalab.org/seurat/articles/integration_introduction.html#performing-integration-on-datasets-normalized-with-sctransform-1).
```{r}
s_list <- SplitObject(s, split.by = 'orig.ident')
s_list <- lapply(s_list, function(s) {
  DefaultAssay(s) <- 'RNA'
  s <- DietSeurat(s, assays = 'RNA')
  s <- SCTransform(s, verbose = FALSE, method = "qpoisson")
  # exclude some genes from variable features: TR, MT, RP
  vf <- VariableFeatures(s)
  exclude <- grepl('^TR[ABGD][VDJ]', vf) | grepl('^MT-', vf) | grepl('^RP[SL]\\d+', vf)
  VariableFeatures(s) <- vf[!exclude]
  return(s)
})

features <- SelectIntegrationFeatures(object.list = s_list, nfeatures = 3000, verbose = FALSE)
s_list <- PrepSCTIntegration(object.list = s_list, anchor.features = features, verbose = FALSE)

anchors <- FindIntegrationAnchors(object.list = s_list, normalization.method = "SCT", 
                                  anchor.features = features, verbose = FALSE)
si <- IntegrateData(anchorset = anchors, normalization.method = "SCT", verbose = FALSE)
si <- RunPCA(si, verbose = FALSE)
dims <- 1:15
si <- RunUMAP(si, reduction = "pca", dims = dims, verbose = FALSE)
si <- FindNeighbors(si, reduction = "pca", dims = dims, verbose = FALSE)
si <- FindClusters(si, resolution = 0.3, verbose = FALSE)


p1 <- DimPlot(si, reduction = "umap", group.by = "orig.ident")
p2 <- DimPlot(si, reduction = "umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE)
```

Show integration and new clustering

```{r, fig.width=10, out.width='100%', fig.height=5}
show(p1 + p2)
```

Show per-cluster composition 
```{r, fig.width=9, out.width='100%', fig.height=3.5}
cp <- composition_plots(df = si@meta.data, group_var = 'seurat_clusters', 
                        comp_var = 'orig.ident', group_name = 'Integration cluster', 
                        comp_name = 'Sample')
show(cp$figure)
```

Show per-cluster QC stats
```{r, fig.width=14, out.width='100%', fig.height=4}
p1 <- VlnPlot(si, features = c('percent.mito', 'log10_nCount_RNA', 'nFeature_RNA'), pt.size = 0.1)
show(p1)
```

Separate samples on UMAP
```{r, out.width='100%', fig.height=3.5, fig.width=14}
p1 <- DimPlot(si, label = TRUE, repel = TRUE, split.by = 'orig.ident') + NoLegend()
show(p1)
```

Show per-sample composition based on integration clusters
```{r, fig.width=9, out.width='100%', fig.height=3.5}
cp <- composition_plots(df = si@meta.data, group_var = 'orig.ident', 
                        comp_var = 'seurat_clusters', group_name = 'Sample', 
                        comp_name = 'Integration cluster', freq_limit = 0)
show(cp$figure & guides(fill=guide_legend(ncol=2)))
```

Look at cluster markers

```{r}
s$SCT@misc$de_res <- diff_mean_test(y = GetAssayData(s, assay = "SCT", slot = "counts"), 
    group_labels = si$seurat_clusters, R = 99, only_pos = TRUE, only_top_n = 333, 
    verbosity = 0)

s$SCT@misc$top_markers <- filter(s$SCT@misc$de_res, mean1 >= 0.2) %>%
    group_by(group1) %>%
    arrange(-mean1, .by_group = TRUE) %>%
    mutate(log2FC = scales::oob_squish_any(x = log2FC, range = c(0, 7)),
           zscore = scales::oob_squish_any(x = zscore, range = c(0, 200)),
           marker_rank = rank(-log2FC*log2FC*zscore)) %>%
    slice_min(order_by = marker_rank, n = 6, with_ties = FALSE) %>% 
    select(group1, gene, mean1, mean2, log2FC, zscore, emp_pval_adj)

# make sure we transfer everything to the merged object for later
s <- GetResidual(s, features = s$SCT@misc$top_markers$gene, verbose = FALSE)
Idents(s) <- si$seurat_clusters
s$integration_cluster <- si$seurat_clusters
s$int_umap <- si$umap
```

```{r, fig.width=10, out.width='100%', fig.height=10}
show(DoHeatmap(s, features = s$SCT@misc$top_markers$gene) + NoLegend())
cells <- WhichCells(s, downsample = 100)
show(DoHeatmap(s, features = s$SCT@misc$top_markers$gene, cells = cells) + NoLegend())
```

Try to find conserved markers?
```{r}
s$SCT@misc$de_res_cons <- diff_mean_test_conserved(
  y = GetAssayData(s, assay = "SCT", slot = "counts"), 
  group_labels = si$seurat_clusters, 
  sample_labels = si$orig.ident, only_pos = TRUE, verbosity = 0)

s$SCT@misc$top_markers_cons <- filter(s$SCT@misc$de_res_cons, de_tests == max(de_tests), log2FC_min > 1, p.adjust(pval_max) <= 0.001) %>% 
  group_by(group1) %>% 
  filter(rank(-log2FC_median, ties.method = "first") <= 6)

s <- GetResidual(s, features = s$SCT@misc$top_markers_cons$gene, verbose = FALSE)
```

```{r, fig.width=10, out.width='100%', fig.height=10}
show(DoHeatmap(s, features = s$SCT@misc$top_markers_cons$gene) + NoLegend())
cells <- WhichCells(s, downsample = 100)
show(DoHeatmap(s, features = s$SCT@misc$top_markers_cons$gene, cells = cells) + NoLegend())
```

Save integration results

```{r}
saveRDS(object = s, file = file.path(rds_dir, '02_integrated_seurat_object.Rds'))
```

Runtime: `r time_diff(start_time)`

Session Info
```{r}
sessionInfo()
```