---
title: 'Azimuth Cell Type Annotation'
author: 'Christoph Hafemeister'
date: '`r format(Sys.time(), "%B %d, %Y %H:%M:%S %Z")`'
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: "hide"
params:
  config_file: "config.yaml"
---

```{r}
# load project-specific parameters
config <- yaml::read_yaml(params$config_file)

suppressPackageStartupMessages({
  library('Matrix')
  library('dplyr')
  library('Seurat')
  library('sctransform')
  library('Azimuth')
  library('reshape2')
  library('ggplot2')
  library('patchwork')
  source(file.path(config$project_root, 'R', 'utils.R'))
})

#future::plan(strategy = "multiprocess", workers = 1)
#options(future.globals.maxSize = 8 * 1024 ^ 3)
#options(future.fork.enable = TRUE)

knitr::opts_chunk$set(comment = NA, fig.width = 5, fig.height = 3.5, out.width = '60%',
                      warning = TRUE, error = TRUE, echo = TRUE, message = TRUE,
                      dpi = 100)
options(dplyr.summarise.inform = FALSE)
options(tibble.width = Inf)
options(DT.options = list(paging = FALSE, scrollX = TRUE, scrollY = '350px',
                          scrollCollapse = TRUE, searching = FALSE, autoWidth = TRUE))

old <- theme_set(theme_bw(base_size = 11))
set.seed(8569205)
start_time <- proc.time()['elapsed']

rds_dir <- file.path(config$out_root, 'Rds')
```

## Load data and predict cell types
```{r}
s <- readRDS(file = file.path(rds_dir, '02_integrated_seurat_object.Rds'))
pred <- azimuth_annotation(query = s, 
                           reference_path = config$azimuth_ref_human_pbmc, 
                           reference_column = 'celltype.l2')
s[['az_umap']] <- pred$umap
s <- AddMetaData(s, pred$prediction)
```

## Visualize the results
```{r, out.width='100%', fig.width=12, fig.height=7}
az_plots <- azimuth_plots(s, reduction = 'az_umap', 
                          annotation = 'az.annotation', 
                          annotation_score = 'az.annotation.score', 
                          mapping_score = 'az.mapping.score')
lapply(az_plots, show)
```

```{r, fig.width=10, fig.height=5, out.width='100%'}
show(DimPlot(s, reduction = "int_umap", group.by = "az.annotation", label = TRUE, repel = TRUE))
#show(DimPlot(s, reduction = "az_umap", group.by = "az.annotation", label = TRUE, repel = TRUE))
```

```{r, fig.width=10, fig.height=10, out.width='100%'}
p1 <- DimPlot(object = s, reduction = "int_umap", split.by = 'az.annotation', ncol = 5) + NoLegend()
show(p1)
```


Show integration-cluster composition based on predictions
```{r, fig.width=9, out.width='90%', fig.height=3.5}
cp <- composition_plots(df = s@meta.data, group_var = 'integration_cluster', 
                        comp_var = 'az.annotation', group_name = 'Integration cluster', 
                        comp_name = 'Cell type', freq_limit = 0.05)
show(cp$figure)
```

Show per-sample composition based on predictions
```{r, fig.width=9, out.width='90%', fig.height=3.5}
cp <- composition_plots(df = s@meta.data, group_var = 'orig.ident', 
                        comp_var = 'az.annotation', group_name = 'Sample', 
                        comp_name = 'Cell type', freq_limit = 0.05)
show(cp$figure)
```


Show fraction details
```{r, out.width='100%', fig.width=10, fig.height=7}
cp <- composition_plots(df = s@meta.data, group_var = 'orig.ident', 
                        comp_var = 'az.annotation', group_name = 'Sample', 
                        comp_name = 'Cell type')
p1 <- ggplot(cp$data, aes(Sample, freq)) + geom_point() +
  ylab('Fraction') + xlab('Sample') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ label_filtered, scales = 'free_y', ncol = 5) +
  theme(legend.position = "none")
show(p1)
```

B cell composition
```{r, fig.width=9, out.width='90%', fig.height=3.5}
cp <- filter(s@meta.data, grepl('^B ', az.annotation)) %>%
  composition_plots(group_var = 'orig.ident', comp_var = 'az.annotation', 
                    group_name = 'Sample', comp_name = 'Cell type')
show(cp$figure)
```


Number of cells
```{r}
tab <- table(s@meta.data$az.annotation, s@meta.data$orig.ident)
print(tab)
```


## Discard low-confidence predictions

Look at composition per sample again,but set predictions where at least one score is lower than 0.5 to NA
```{r}
s$az.annot.filtered <- s$az.annotation
sel <- s$az.annotation.score >= 0.5 & s$az.mapping.score >= 0.5
s$az.annot.filtered[!sel] <- NA

```

Show integration-cluster composition based on predictions
```{r, fig.width=9, out.width='90%', fig.height=3.5}
cp <- composition_plots(df = s@meta.data, group_var = 'integration_cluster', 
                        comp_var = 'az.annot.filtered', group_name = 'Integration cluster', 
                        comp_name = 'Cell type', freq_limit = 0.05)
show(cp$figure)
```

Show per-sample composition based on predictions
```{r, fig.width=9, out.width='90%', fig.height=3.5}
cp <- composition_plots(df = s@meta.data, group_var = 'orig.ident', 
                        comp_var = 'az.annot.filtered', group_name = 'Sample', 
                        comp_name = 'Cell type', freq_limit = 0.05)
show(cp$figure)
```

Remove NA-predicted cells
```{r, fig.width=9, out.width='90%', fig.height=3.5}
cp <- filter(s@meta.data, !is.na(az.annot.filtered)) %>% 
  composition_plots(group_var = 'orig.ident', comp_var = 'az.annot.filtered', 
                    group_name = 'Sample', comp_name = 'Cell type', 
                    freq_limit = 0.05)
show(cp$figure)
```


Show fraction details
```{r, out.width='100%', fig.width=10, fig.height=7}
cp <- filter(s@meta.data, !is.na(az.annot.filtered)) %>% 
  composition_plots(group_var = 'orig.ident', comp_var = 'az.annot.filtered', 
                    group_name = 'Sample', comp_name = 'Cell type')
p1 <- ggplot(cp$data, aes(Sample, freq)) + geom_point() +
  ylab('Fraction') + xlab('Sample') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ label_filtered, scales = 'free_y', ncol = 5) +
  theme(legend.position = "none")
show(p1)
```


## Save, Runtime, Session info

```{r}
saveRDS(object = s@meta.data, file = file.path(rds_dir, '03_integrated_annotated_meta_data.Rds'))
saveRDS(object = s$az_umap, file = file.path(rds_dir, '03_integrated_annotated_az_umap.Rds'))
```

Runtime: `r time_diff(start_time)`

Session Info
```{r}
sessionInfo()
```
