---
title: 'Immune profiling analysis'
author: 'Christoph Hafemeister'
date: '`r format(Sys.time(), "%B %d, %Y %H:%M:%S %Z")`'
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    toc_float: TRUE
    code_folding: "hide"
params:
  config_file: "config.yaml"
---

```{r}
# load project-specific parameters
config <- yaml::read_yaml(params$config_file)

suppressPackageStartupMessages({
  library('Matrix')
  library('dplyr')
  library('Seurat')
  library('sctransform')
  library('reshape2')
  library('ggplot2')
  library('patchwork')
  library('ggrepel')
  source(file.path(config$project_root, 'R', 'utils.R'))
})

#future::plan(strategy = "multiprocess", workers = 1)
#options(future.globals.maxSize = 8 * 1024 ^ 3)
#options(future.fork.enable = TRUE)

knitr::opts_chunk$set(comment = NA, fig.width = 5, fig.height = 3.5, out.width = '60%',
                      warning = TRUE, error = TRUE, echo = TRUE, message = TRUE,
                      dpi = 100)
options(dplyr.summarise.inform = FALSE)
options(tibble.width = Inf)
options(DT.options = list(paging = TRUE, searching = FALSE, autoWidth = TRUE))

old <- theme_set(theme_bw(base_size = 11))
set.seed(8569205)
start_time <- proc.time()['elapsed']

rds_dir <- file.path(config$out_root, 'Rds')
```

# Load scRNA-seq results
```{r}
# load results of cell type annotation 
s_md <- readRDS(file = file.path(rds_dir, '03_integrated_annotated_meta_data.rds'))
s_md$barcode <- rownames(s_md)
```


# Immune profiling analysis

## Read data; first overview

Read clonotype contig annotations into one big data frame and print table of Sample ID vs Chain detected. Note that we keep only observations with productive contig (likely functional T or B cell receptor) and annotated with a single chain (discarding 'Multi' chains that we see in gdTCR results quite often)
```{r}
path = config$data_root
clonotype_files <- list.files(path = path, pattern = '^filtered_contig_annotations\\.csv$', recursive = TRUE, full.names = TRUE)
ret <- list()
for (f in clonotype_files) {
  clono_dat <- read.csv(f)
  clono_dat$dir_name <- basename(dirname(f))
  # also read the clonotype file, so we can add the iNKT and MAIT evidence
  clono_anno <- read.csv(file = gsub(pattern = 'filtered_contig_annotations\\.csv$', 
                                     x = f,
                                     replacement = 'clonotypes.csv'))
  if (all(c('clonotype_id', 'inkt_evidence', 'mait_evidence') %in% colnames(clono_anno))) {
    clono_dat <- left_join(clono_dat, clono_anno[, c('clonotype_id', 'inkt_evidence', 'mait_evidence')], by = c('raw_clonotype_id' = 'clonotype_id'))
  } else {
    clono_dat$inkt_evidence <- ''
    clono_dat$mait_evidence <- ''
  }
  
  ret[[basename(dirname(f))]] <- clono_dat
}
clono_dat <- do.call(rbind, ret)
#print(table(clono_dat$dir_name, clono_dat$productive))
clono_dat <- do.call(rbind, ret) %>% filter(productive %in% c('true', 'True'), chain != 'Multi')
```

Table of directory name and chain
```{r}
print(table(clono_dat$dir_name, clono_dat$chain))
```

Infer sample IDs from directory names and show sample-chain table
```{r}
# infer sample ids from directory names
clono_dat$sample_id <- gsub(pattern = '_(abTCR|BCR|gdTCR)_VDJ(_GD)*', replacement = '', x = clono_dat$dir_name)
print(table(clono_dat$sample_id, clono_dat$chain))

clono_dat$sample_number <- as.integer(stringr::str_extract(clono_dat$sample_id, '\\d+'))
# make sure the cell barcode matches the names assigned by Seurat in the merged and integrated analysis
clono_dat$barcode <- paste(clono_dat$barcode, clono_dat$sample_number, sep = '_')
```

Aggregate the chain counts per cell. How often do we see certain chains together (and how many of them)?
```{r}
chain_counts <- arrange(clono_dat, sample_id, barcode, chain) %>%
  group_by(sample_id, barcode, chain) %>% 
  summarise(n = n()) %>%
  mutate(chain_count = paste(chain, n, sep=':')) %>%
  summarise(chains = paste(chain_count, collapse = ';'))
tab <- as.data.frame(sort(table(chain_counts$chains), decreasing = TRUE))
colnames(tab) <- c('Chain counts', 'Cells')
DT::datatable(tab)
```


```{r}
# combine the information per cell; use only data with productive CDR3
clono_sum <- arrange(clono_dat, sample_id, barcode, chain) %>%
  group_by(sample_id, barcode) %>% 
  mutate(chain_cdr3 = paste(chain, cdr3, sep=':')) %>%
  summarise(clonotype = paste(chain_cdr3, collapse = ';'),
            inkt_evidence = paste(setdiff(sort(unique(inkt_evidence)), ''), collapse = ';'),
            mait_evidence = paste(setdiff(sort(unique(mait_evidence)), ''), collapse = ';')) %>%
  mutate(has_TRAB = grepl(pattern = 'TR[AB]:', x = clonotype),
         has_IG = grepl(pattern = 'IG[HKL]:', x = clonotype),
         has_TRDG = grepl(pattern = 'TR[DG]:', x = clonotype),
         only_TRAB = has_TRAB & !has_IG & !has_TRDG,
         only_IG = !has_TRAB & has_IG & !has_TRDG,
         only_TRDG = !has_TRAB & !has_IG & has_TRDG)

freq_ab <- filter(clono_sum, has_TRAB) %>%
  group_by(sample_id, clonotype) %>%
  summarise(n = n()) %>%
  mutate(freq = n / sum(n), rank = rank(-freq, ties.method = 'min')) %>% 
  arrange(sample_id, -freq)

freq_ig <- filter(clono_sum, has_IG) %>%
  group_by(sample_id, clonotype) %>%
  summarise(n = n()) %>%
  mutate(freq = n / sum(n), rank = rank(-freq, ties.method = 'min')) %>% 
  arrange(sample_id, -freq)

freq_dg <- filter(clono_sum, has_TRDG) %>%
  group_by(sample_id, clonotype) %>%
  summarise(n = n()) %>%
  mutate(freq = n / sum(n), rank = rank(-freq, ties.method = 'min')) %>% 
  arrange(sample_id, -freq)

# join with scRNA-seq results
s_md <- left_join(s_md, clono_sum, by = 'barcode')
pred_cell_types <- sort(unique(as.character(s_md$az.annotation)))
s_md$az.annotation <- factor(s_md$az.annotation, levels = pred_cell_types)
s_md <- left_join(s_md, chain_counts)
```

Save for later
```{r}
saveRDS(object = list(clono_dat = clono_dat, 
                      clono_sum = clono_sum,
                      md = s_md), 
        file = file.path(rds_dir, '04_clonotypes.Rds'))
```


## Runtime and session info

Runtime: `r time_diff(start_time)`

Session Info
```{r}
sessionInfo()
```
